---
title: "Project_BI"
author: "Petrenko Kate"
date: '2023-01-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("dplyr", quietly = TRUE)) BiocManager::install("dplyr")
library(dplyr)
if (!requireNamespace("tidyr", quietly = TRUE)) BiocManager::install("tidyr")
library(tidyr)
if (!requireNamespace("ggplot2", quietly = TRUE)) BiocManager::install("ggplot2")
library(ggplot2)
if (!requireNamespace("ggbreak", quietly = TRUE)) BiocManager::install("ggbreak")
library(ggbreak)
if (!requireNamespace("stringr", quietly = TRUE)) BiocManager::install("stringr")
library(stringr)
if (!requireNamespace("data.table", quietly = TRUE)) BiocManager::install("data.table")
library(data.table)
if (!requireNamespace("readr", quietly = TRUE)) BiocManager::install("readr")
library(readr)
if (!requireNamespace("clusterProfiler", quietly = TRUE)) BiocManager::install("clusterProfiler")
library(clusterProfiler)
if (!requireNamespace("org.Sc.sgd.db", quietly = TRUE)) BiocManager::install("org.Sc.sgd.dbb")
library(org.Sc.sgd.db)
if (!requireNamespace("parallel", quietly = TRUE)) BiocManager::install("parallel")
library(parallel)
```

# Pre-processing
## Read data

```{r}
#PPI raw data
biogrid <- read.csv('data/BIOGRID-ORGANISM-Saccharomyces_cerevisiae_S288c-4.4.217.tab3.txt', sep='\t') %>% 
  as.data.table()

#GI PSN
costanzo <- read.csv('data/PSN/cc_ALL.txt', sep = '\t', skip = 1)

#GI raw data
SGA_ExE <- fread('../../Downloads/Data File S1. Raw genetic interaction datasets- Pair-wise interaction format/SGA_ExE.txt',sep = '\t') #df1
SGA_ExN <- fread('../../Downloads/Data File S1. Raw genetic interaction datasets- Pair-wise interaction format/SGA_ExN_NxE.txt', sep='\t') #df2
SGA_NxN <- fread('../../Downloads/Data File S1. Raw genetic interaction datasets- Pair-wise interaction format/SGA_NxN.txt', sep='\t', header = T) #df3
```

## Transformation of data table
### GI PSN
```{r}
costanzo_gene_names <- costanzo[, c(1,2)]
costanzo <- costanzo[-1]

costanzo <- costanzo %>% 
  gather(key = 'gene_B', value = 'Score', -X.1) %>% 
  dplyr::filter(Score != 'NaN') %>% 
  as.data.table()

names(costanzo)[1] <- 'gene_A'

head(costanzo)
```

### PPI
```{r}
ppi_stats <- biogrid %>% 
  dplyr::filter(Experimental.System.Type == 'physical') %>% 
  dplyr::count(Experimental.System)

na_expsystem <- biogrid %>% 
  group_by(Experimental.System) %>% 
  summarise(non_na = sum(Score != '-')) %>% 
  dplyr::filter(Experimental.System %in% ppi_stats$Experimental.System)

ppi_stats <- merge(ppi_stats, na_expsystem) %>% 
  mutate(non_na = non_na / n * 100)

ppi_stats

PPI <- biogrid %>% 
  filter(Experimental.System.Type == 'physical') %>% 
  unique()
```

### GI raw

```{r}
SGA <- bind_rows(SGA_ExE[,c(1,3,4,6,7)], SGA_ExN[,c(1,3,4,6,7)], SGA_NxN[,c(1,3,4,6,7)], .id = 'df') 

SGA$`Query Strain ID` <- 
  SGA$`Query Strain ID` %>% 
  gsub('_.*', '', .)

SGA$`Array Strain ID` <- 
  SGA$`Array Strain ID` %>% 
  gsub('_.*', '', .)

names(SGA)[c(2,3)] <- c('Systematic.Name.Interactor.A', 'Systematic.Name.Interactor.B')

head(SGA)
```

## Cutoff 
### Score distribution in the GI (PSN)
```{r}
costanzo %>% 
  ggplot(aes(Score)) +
  geom_histogram(binwidth = 0.01, color='black', fill='lightblue') +
  scale_x_continuous(breaks = seq(-1, 1, 0.2)) +
  theme_bw()
```

## Cutoff for GI (PSN) = 0.2
```{r}
costanzo_02 <- costanzo %>% 
  dplyr::filter(Score >= 0.2) %>% 
  unique() %>% 
  arrange(desc(Score))

names(costanzo_02) <- c('Systematic.Name.Interactor.A', 'Systematic.Name.Interactor.B', 'Score')

mask.dup <-
  paste(costanzo_02$Systematic.Name.Interactor.A, costanzo_02$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated
sum(!mask.dup)

nrow(costanzo_02)

costanzo_02_uniq <- costanzo_02[!mask.dup,]
```

### Common patterns between GI (PSN, cutoff = 0.2) and PPI 
```{r}
common_GI_PPI_02 <- inner_join(costanzo_02, PPI[,c(6,7,12, 19)], 
                               PPI[,c(6,7,12,19)] %>% 
                                 dplyr::mutate(A = Systematic.Name.Interactor.B,
                                               B = Systematic.Name.Interactor.A,
                                               Systematic.Name.Interactor.A = A,
                                               Systematic.Name.Interactor.B = B),   
                                 by=c('Systematic.Name.Interactor.A', 
                                      'Systematic.Name.Interactor.B')) %>% 
  unique() %>%
  dplyr::arrange(desc(Score.x))
  
mask.dup <-
  paste(common_GI_PPI_02$Systematic.Name.Interactor.A, common_GI_PPI_02$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated

common_GI_PPI_02 <- common_GI_PPI_02[!mask.dup,]

names(common_GI_PPI_02)[c(3, 5)] <- c('Score_GI', 'Score_PPI') 
nrow(common_GI_PPI_02)
```

## Cutoff for GI (raw)

```{r}
SGA_intermediate_1 <- SGA %>% 
  dplyr::filter(`Genetic interaction score (ε)` > 0.08 & `P-value` < 0.05) %>% 
  unique() %>% 
  arrange(desc(`Genetic interaction score (ε)`))

SGA_intermediate_2 <- SGA %>% 
  dplyr::filter(`Genetic interaction score (ε)` < -0.08 & `P-value` < 0.05) %>% 
  unique() %>% 
  arrange(`Genetic interaction score (ε)`)

SGA_intermediate <- bind_rows(SGA_intermediate_1, SGA_intermediate_2) %>% 
  unique()
rm(SGA_intermediate_1, SGA_intermediate_2)

SGA_stringent_1 <- SGA %>% 
  dplyr::filter(`P-value` < 0.05 & `Genetic interaction score (ε)` > 0.16) %>% 
  unique() %>% 
  arrange(desc(`Genetic interaction score (ε)`))
  
SGA_stringent_2 <- SGA %>% 
  dplyr::filter(`P-value` < 0.05 & `Genetic interaction score (ε)` < -0.12) %>% 
  unique() %>% 
  arrange(`Genetic interaction score (ε)`)

SGA_stringent <- bind_rows(SGA_stringent_1, SGA_stringent_2) %>% 
  unique()
rm(SGA_stringent_1, SGA_stringent_2)

SGA_intermediate[,2:3] %>% 
  unique() %>% 
  nrow()

mask.dup <-
  paste(SGA_intermediate$Systematic.Name.Interactor.A, SGA_intermediate$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated
sum(!mask.dup)

SGA_stringent[,2:3] %>% 
  unique() %>% 
  nrow()

mask.dup <-
  paste(SGA_stringent$Systematic.Name.Interactor.A, SGA_stringent$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated
sum(!mask.dup)
```

### Common patterns between GI (raw, intermediate) and PPI

```{r}
common_GI_PPI_SGA_inter <- inner_join(SGA_intermediate, PPI[,c(6,7,12, 19)], by=c('Systematic.Name.Interactor.A', 'Systematic.Name.Interactor.B')) %>% 
  unique()

names(common_GI_PPI_SGA_inter)[8] <- 'Score_PPI' 

mask.dup <-
  paste(common_GI_PPI_SGA_inter$Systematic.Name.Interactor.A, common_GI_PPI_SGA_inter$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated

common_GI_PPI_SGA_inter <- common_GI_PPI_SGA_inter[!mask.dup,]
```

### Common patterns between GI (raw, stringent) and PPI

```{r}
common_GI_PPI_SGA_strong <- inner_join(SGA_stringent, PPI[,c(6,7,12, 19)], by=c('Systematic.Name.Interactor.A', 'Systematic.Name.Interactor.B')) %>% 
  unique()
names(common_GI_PPI_SGA_strong)[8] <- 'Score_PPI' 

mask.dup <-
  paste(common_GI_PPI_SGA_strong$Systematic.Name.Interactor.A, common_GI_PPI_SGA_strong$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated

common_GI_PPI_SGA_strong <- common_GI_PPI_SGA_strong[!mask.dup,]
```

# Gene-level table

```{r}
union_PSN_02 <- dplyr::union(costanzo_02[,c(1,2)], PPI[, c(6, 7)])

mask.dup <-
  paste(union_PSN_02$Systematic.Name.Interactor.A, union_PSN_02$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated

union_PSN_02 <- union_PSN_02[!mask.dup, ]

union_SGA_inter <- dplyr::union(SGA_intermediate[,c(2,3)], PPI[, c(6, 7)])

mask.dup <-
  paste(union_SGA_inter$Systematic.Name.Interactor.A, union_SGA_inter$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated

union_SGA_inter <- union_SGA_inter[!mask.dup, ]
```

OI (intersection/smallest set size) 
JI (intersection/union)

```{r}
gene_level <- unique(costanzo_gene_names$X.1)  %>%
  as.data.table()
  
names(gene_level) <- 'Gene'

gene_level <- gene_level  %>% 
  mutate(OI_PSN_02 = sapply(Gene, function(x)
                             ((sum(common_GI_PPI_02$Systematic.Name.Interactor.A == x) +
                             sum(common_GI_PPI_02$Systematic.Name.Interactor.B == x)) / 
                               (sum(PPI$Systematic.Name.Interactor.A == x) +
                                  sum(PPI$Systematic.Name.Interactor.B  == x))))) %>% 
  mutate(JI_PSN_02 = sapply(Gene, function(x)
                             ((sum(common_GI_PPI_02$Systematic.Name.Interactor.A == x) +
                             sum(common_GI_PPI_02$Systematic.Name.Interactor.B == x)) / 
                               (sum(union_PSN_02$Systematic.Name.Interactor.A == x) +
                                  sum(union_PSN_02$Systematic.Name.Interactor.B == x))))) %>% 
  mutate(OI_SGA_inter = sapply(Gene, function(x)
                             ((sum(common_GI_PPI_SGA_inter$Systematic.Name.Interactor.A == x) +
                             sum(common_GI_PPI_SGA_inter$Systematic.Name.Interactor.B == x)) / 
                               (sum(PPI$Systematic.Name.Interactor.A == x) +
                                  sum(PPI$Systematic.Name.Interactor.B  == x))))) %>% 
  mutate(JI_SGA_inter = sapply(Gene, function(x)
                             ((sum(common_GI_PPI_SGA_inter$Systematic.Name.Interactor.A == x) +
                             sum(common_GI_PPI_SGA_inter$Systematic.Name.Interactor.B == x)) / 
                               (sum(union_SGA_inter$Systematic.Name.Interactor.A == x) +
                                  sum(union_SGA_inter$Systematic.Name.Interactor.B == x)))))

```

## Visualisation of gene-level

```{r, figures-side, fig.show="hold", out.width="50%"}

ggplot(gene_level, aes(x=OI_PSN_02*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("OI between PSN (cutofff = 0.2) and PPI") +
  theme_light() +
  theme(plot.title = element_text(size=15))

gene_level %>% 
  filter(OI_PSN_02 != 0) %>% 
  ggplot( aes(x=OI_PSN_02*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("OI between PSN (cutofff = 0.2) and PPI without 0") +
  theme_light() +
  theme(plot.title = element_text(size=15))

ggplot(gene_level, aes(x=JI_PSN_02*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("JI between PSN (cutofff = 0.2) and PPI") +
  theme_light() +
  theme(plot.title = element_text(size=15))

gene_level %>% 
  filter(JI_PSN_02 != 0) %>% 
  ggplot( aes(x=JI_PSN_02*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("JI between PSN (cutofff = 0.2) and PPI without 0") +
  theme_light() +
  theme(plot.title = element_text(size=15))

ggplot(gene_level, aes(x=OI_SGA_inter*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("OI between SGA (intermediate confidence) and PPI") +
  theme_light() +
  theme(plot.title = element_text(size=15))

gene_level %>% 
  dplyr::filter(OI_SGA_inter != 0) %>% 
  ggplot( aes(x=OI_SGA_inter*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("OI between SGA (intermediate confidence) and PPI without 0") +
  theme_light() +
  theme(plot.title = element_text(size=15))

ggplot(gene_level, aes(x=JI_SGA_inter*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("JI between SGA (intermediate confidence) and PPI") +
  theme_light() +
  theme(plot.title = element_text(size=15))

gene_level %>% 
  dplyr::filter(JI_SGA_inter != 0) %>% 
  ggplot( aes(x=JI_SGA_inter*100)) +
  geom_histogram(binwidth = 1, color='#333233', fill='#9779c7', alpha=0.8) +
  ggtitle("JI between SGA (intermediate confidence) and PPI without 0") +
  theme_light() +
  theme(plot.title = element_text(size=15))
```

## Hex-plots to estimate the correlation between the number of interactions and JI
### Pre-processing data
```{r}
PSN_02_JI_cor <- unique(costanzo_gene_names$X.1)  %>%
  as.data.table()

names(PSN_02_JI_cor) <- 'Gene'

mask.dup <- paste(PPI$Systematic.Name.Interactor.A, PPI$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated

PPI_uniq <- PPI[!mask.dup, ]

PSN_02_JI_cor <- PSN_02_JI_cor %>% 
  mutate(JI = sapply(Gene, function(x)
    ((sum(common_GI_PPI_02$Systematic.Name.Interactor.A == x) +
        sum(common_GI_PPI_02$Systematic.Name.Interactor.B == x)) / 
       (sum(union_PSN_02$Systematic.Name.Interactor.A == x) +
          sum(union_PSN_02$Systematic.Name.Interactor.B == x))))) %>% 
  mutate(PPI_n = sapply(Gene, function(x)
    (sum(PPI_uniq$Systematic.Name.Interactor.A == x) + 
       sum(PPI_uniq$Systematic.Name.Interactor.B == x)))) %>% 
  mutate(GI_n = sapply(Gene, function(x)
    (sum(costanzo_02_uniq$Systematic.Name.Interactor.A == x) + 
       sum(costanzo_02_uniq$Systematic.Name.Interactor.B == x))))
```

### Plots

```{r}
ggplot(PSN_02_JI_cor, aes(x=JI, y=PPI_n) ) +
  geom_hex() +
  theme_light() 

ggplot(PSN_02_JI_cor, aes(x=JI, y=GI_n) ) +
  geom_hex() +
  theme_light()
```

# Enrichment analysis
lowest (Jaccard = 0 AND PPI_n >= 5 AND GI_n >= 5)
```{r}
annotate_data <- PSN_02_JI_cor %>% 
  dplyr::arrange(desc(JI)) %>% 
  dplyr::slice(1:100)

annotate_data <- PSN_02_JI_cor %>% 
  dplyr::filter(JI == 0) %>%
  dplyr::filter(PPI_n >= 5) %>% 
  dplyr::filter(GI_n >= 5) %>% 
  rbind(annotate_data) %>% 
  mutate(ENTREZID =  mapIds(org.Sc.sgd.db, 
                            keys = Gene, 
                            column = "ENTREZID", 
                            keytype = "ORF"))

annotate_data_highest<- annotate_data %>% 
  dplyr::filter(JI != 0)
 
annotate_data_lowest <- annotate_data %>% 
  dplyr::filter(JI == 0) %>% 
  dplyr::filter(PPI_n >= 23) %>% 
  dplyr::filter(GI_n >= 23)

annotate_data_23 <- PSN_02_JI_cor %>% 
  dplyr::filter(PPI_n >= 23) %>% 
  dplyr::filter(GI_n >= 23) %>% 
  mutate(ENTREZID =  mapIds(org.Sc.sgd.db, 
                            keys = Gene, 
                            column = "ENTREZID", 
                            keytype = "ORF")) 

ego23 <- enrichGO(annotate_data$ENTREZID, org.Sc.sgd.db, ont = "ALL")
  
ego <- enrichGO(annotate_data$ENTREZID, org.Sc.sgd.db, ont = "ALL")

ego_uni <- enrichGO(annotate_data_lowest$ENTREZID, org.Sc.sgd.db, 
                    ont = "ALL", universe = annotate_data_23$ENTREZID)

ego_highest <- enrichGO(annotate_data_highest$ENTREZID, org.Sc.sgd.db, ont = "ALL")

ego_lowest <- enrichGO(annotate_data_lowest$ENTREZID, org.Sc.sgd.db, ont = "ALL")
```

```{r  out.width = "100%"}
dotplot(ego_uni, split="ONTOLOGY", showCategory = 15, font.size = 8) + 
  facet_grid(ONTOLOGY~., scale="free")

barplot(ego_uni, showCategory = 15, font.size = 8)

dotplot(ego23, showCategory = 15, font.size = 8)
```

## Plots for highest

```{r out.width = "100%"}
dotplot(ego_highest, split="ONTOLOGY", showCategory = 15) + 
  facet_grid(ONTOLOGY~., scale="free") 

barplot(ego_highest)
```


## Plots for lowest

```{r out.width = "100%"}
dotplot(ego_lowest, split="ONTOLOGY", showCategory = 15) + 
  facet_grid(ONTOLOGY~., scale="free") 

barplot(ego_lowest)
```

## Plots for top-100 genes by JI after filter n

Try running enrichment analysis for genes with highest JI by first taking genes with (GI_n >= AND PPI_n >= some cutoff). After filtering by GI_n and PPI_n, take top-100 genes by JI and show their lowest JI value (i.e., JI for the gene ranked 100th);

```{r}
annotate_data_topgenes <- PSN_02_JI_cor %>% 
  dplyr::filter(PPI_n >= 23) %>% 
  dplyr::filter(GI_n >= 23) %>%
  dplyr::arrange(desc(JI)) %>%
  dplyr::slice(1:100) %>% 
  mutate(ENTREZID =  mapIds(org.Sc.sgd.db, 
                            keys = Gene, 
                            column = "ENTREZID", 
                            keytype = "ORF"))

annotate_data_all <- PSN_02_JI_cor %>% 
  mutate(ENTREZID =  mapIds(org.Sc.sgd.db, 
                            keys = Gene, 
                            column = "ENTREZID", 
                            keytype = "ORF"))
  

ego_23 <- enrichGO(annotate_data_topgenes$ENTREZID, org.Sc.sgd.db, 
                   ont = "ALL", universe = annotate_data_23$ENTREZID)

ego_all <- enrichGO(annotate_data_topgenes$ENTREZID, org.Sc.sgd.db, 
                    ont = "ALL", universe = annotate_data_all$ENTREZID)
```

plots for universe 23
```{r}
dotplot(ego_23, split="ONTOLOGY", showCategory = 15) + 
  facet_grid(ONTOLOGY~., scale="free") 

barplot(ego_23)
```
plots for universe all
```{r}
dotplot(ego_all, split="ONTOLOGY", showCategory = 15) + 
  facet_grid(ONTOLOGY~., scale="free") 

barplot(ego_all)
```

# Destribution of interaction number

```{r}
top100JI <- PSN_02_JI_cor %>% 
      dplyr::arrange(desc(JI)) %>% 
      dplyr::slice(1:100)

destr_inter <- PSN_02_JI_cor %>% 
  dplyr::arrange(desc(JI)) %>% 
  dplyr::select(Gene, GI_n, PPI_n) %>%  
  dplyr::mutate(GI_n_topJI = ifelse(
    Gene %in% top100JI$Gene, GI_n, NA), 
    PPI_n_topJI = ifelse(
      Gene %in% top100JI$Gene, PPI_n, NA)
    ) %>% 
  pivot_longer(cols=c('GI_n', 'PPI_n', 
                      'GI_n_topJI', 'PPI_n_topJI')) %>% 
  dplyr::filter(value < 200) %>% 
  dplyr::rename(n = value)
  
ggplot(destr_inter, aes(x=name, y=n)) +
    geom_violin(width=1.4) +
    geom_boxplot(width=0.1, alpha=0.2) +
  theme_bw()
```

# PSN for PPI

```{r}

#JI calculation in PPI_uniq only
#x - gene names after paste(sep = ',')
ji_calc <- function(x) {
  
  a = str_split(x, ',')[[1]][1]
  b = str_split(x, ',')[[1]][2]
  
  a_inter <- PPI_uniq %>% 
    dplyr::select(Systematic.Name.Interactor.A,
                  Systematic.Name.Interactor.B) %>% 
    dplyr::filter(Systematic.Name.Interactor.A == a | 
                  Systematic.Name.Interactor.B == a) %>% 
    as.matrix() %>% 
    as.vector() 
  
  a_inter <- a_inter[!a_inter == a]  
  
  b_inter <- PPI_uniq %>% 
    dplyr::select(Systematic.Name.Interactor.A,
                  Systematic.Name.Interactor.B) %>% 
    dplyr::filter(Systematic.Name.Interactor.A == b | 
                  Systematic.Name.Interactor.B == b) %>% 
    as.matrix() %>% 
    as.vector()
  
  b_inter <- b_inter[!b_inter == b]

  length(intersect(a_inter, b_inter))/length(union(a_inter, b_inter))

}

ppi_ji <- PPI_uniq %>% 
  dplyr::select(Systematic.Name.Interactor.A,
                Systematic.Name.Interactor.B) %>%
  dplyr::rename(Gene.A=Systematic.Name.Interactor.A,
                Gene.B = Systematic.Name.Interactor.B) %>%  
  dplyr::mutate(JI = mclapply(paste(Gene.A, Gene.B, sep = ','), 
                              ji_calc, mc.cores = 4) %>% unlist)

ppi_ji <- ppi_ji %>% 
  dplyr::filter(Gene.A != Gene.B) %>% 
  dplyr::mutate(JI=unlist(JI))

n_calc <- function(x) {
  
      sum(PPI_uniq$Systematic.Name.Interactor.A == x | 
          PPI_uniq$Systematic.Name.Interactor.B == x) 
    
}

ppi_ji <- ppi_ji %>% 
  dplyr::mutate(B.n = mclapply(Gene.B, n_calc, mc.cores = 4) %>% unlist,
                A.n = mclapply(Gene.A, n_calc, mc.cores = 4) %>% unlist)
  
ppi_ji %>% 
  pivot_longer(cols=c('A.n', 'B.n')) %>% 
  ggplot(aes(x=name, y=value)) +
    geom_violin(width=1.4) +
    geom_boxplot(width=0.1, alpha=0.2) +
  theme_bw()
```


## Density JI in PPI

```{r}
hist(ppi_ji$JI, labels = T, border = F, main = '', col=rgb(0.2,0.8,0.5,0.5))
ppi_ji %>% 
  ggplot(aes(x=JI)) +
  geom_histogram(binwidth = 0.01, color='black', fill='lightblue') +
  scale_y_log10() +
  theme_bw()

```

## Common PPI PSN and GI PSN
```{r}
common_GI_PPI_PSN <- inner_join(costanzo_02, ppi_ji, by=c('Systematic.Name.Interactor.A', 'Systematic.Name.Interactor.B')) %>% 
  unique() %>%
  dplyr::arrange(desc(Score.x))

mask.dup <-
  paste(common_GI_PPI_PSN$Systematic.Name.Interactor.A, common_GI_PPI_PSN$Systematic.Name.Interactor.B, sep = ',') %>%
  str_split(',') %>% 
  lapply(function(x) x %>% 
           sort %>% 
           paste(collapse = ',')) %>%
  unlist %>%
  duplicated

common_GI_PPI_PSN <- common_GI_PPI_PSN[!mask.dup,]

names(common_GI_PPI_PSN)[c(3, 5)] <- c('Score_GI', 'Score_PPI') 
nrow(common_GI_PPI_PSN)
```

## Enrichment 
```{r}
annotate_data <- PSN_02_JI_cor %>% 
  dplyr::filter(PPI_n >= 23) %>% 
  dplyr::filter(GI_n >= 23) %>% 
  dplyr::arrange(desc(JI)) %>% 
  dplyr::slice(1:100) %>% 
  mutate(ENTREZID =  mapIds(org.Sc.sgd.db, 
                            keys = Gene, 
                            column = "ENTREZID", 
                            keytype = "ORF"))

annotate_data_uni <- PSN_02_JI_cor %>% 
  mutate(ENTREZID =  mapIds(org.Sc.sgd.db, 
                            keys = Gene, 
                            column = "ENTREZID", 
                            keytype = "ORF")) 

ego <- enrichGO(annotate_data$ENTREZID, org.Sc.sgd.db, ont = "ALL", universe = annotate_data_uni$ENTREZID)

dotplot(ego, split="ONTOLOGY", showCategory = 8, font.size=9) + 
  facet_grid(ONTOLOGY~., scale="free") 

barplot(ego)
```
```{r}
ego <- enrichGO(annotate_data$ENTREZID, org.Sc.sgd.db, ont = "ALL", universe = annotate_data_23$ENTREZID)

dotplot(ego, split="ONTOLOGY", showCategory = 8, font.size=9) + 
  facet_grid(ONTOLOGY~., scale="free") 

barplot(ego)
```

